\section{Specification}
In pharmacokinetics, modelling approaches for drug distribution based on ordinary differential equations fall short in light of intricate phenomena involving anomalous diffusion, deep tissue trapping, long-term accumulation effects, diffusion across capillaries, synergistic and competitive action and many more. These give rise to fractional-order pharmacokinetics.

Fractional-order dynamic equations rely on notions of fractional-order derivatives (e.g. derivatives of order 0.5); several definitions of fractional-order derivatives have been proposed in the literature. Such derivatives are used to formulate fractional-order dynamical systems which are suitable for modelling the distribution of compounds in the body.

The focus of this project will be on the development of fractional-order models for chemo-therapeutics against ovarian cancer. Relevant data have been provided by Prof Yahya Choonara (Dept of Pharmacology, University of Witwatersrand, Johannesburg, South Africa), with whom we are involved in a collaborative research project. That said, this work can have significant implications for drug discovery.\newline

\textbf{Objectives}
\begin{enumerate}
  \item Understand the principles of fractional-order derivatives and their calculus and compartmental pharmacokinetic models
  \item Employ appropriate statistical estimation methods to build fractional-order models from data
\end{enumerate}
\textbf{MEng Extension}\newline \newline
At the MEng level, the student will focus on the development of an estimation methodology for the pharmacokinetic parameter of fractional-order models from available data. The focus will be on maximum likelihood estimation methods and maximum a posteriori methods.\newline

\newpage
\textbf{Learning Outcomes}\newline

Upon completion of the project the student should expect:

\begin{enumerate}
  \item To have gained a thorough understanding of fractional calculus and
  \item A solid understanding of the basic principles of pharmacokinetics (including compartmental modelling)
  \item To have developed fractional-order pharmacokinetic models for anti-ovarian cancer compounds
\end{enumerate}
